Experience Report: Haskell in Computational Biology

### 2012-03-08 mrfy_experience_report.tex
added a barriers section

%%  \subsection{Our Background}
%%  
%%  All three authors are members of the Tufts University Department of Computer 
%%  Science. Noah Daniels has taken a graduate seminar in advanced functional 
%%  programming, which included some Haskell, and spent ten years in industry as a 
%%  professional programmer in languages such as Ruby, C, and C++. Andrew Gallant 
%%  has taken a programming languages course heavily focused on functional 
%%  programming, but with no Haskell. Norman Ramsey is the local expert, but he 
%%  wrote no code for this project.

@@

\subsection{Barriers}

- profiling
- debugging
 [ talk about profiling and debugging problems where?]
[ something about quickcheck: how do we know when we have enough QC props? How do we start testing a complex module? ]

>>>>

@@

\subsection{Our Background}

\remark{I moved this section from the Intro as per NR's suggestion. ---AG}
All three authors are members of the Tufts University Department of Computer
Science.
Noah Daniels has taken a graduate seminar in advanced functional
programming, which included some Haskell, and spent ten years in industry as a
professional programmer in languages such as Ruby, C, and C++.
Andrew Gallant has taken a programming languages course heavily focused on 
functional programming, but with no Haskell.
Norman Ramsey is the local expert, but he wrote no code for this project.

\subsection{Barriers}
\remark{I feel like a lot of this is a summary of some of the problems we've
  discussed previously in the paper. Is that okay? A summary may be worth it. 
---AG}

As inexperienced functional programmers, developing a complex piece of
software in Haskell posed several barriers.
\remark{NOAH: You did the profiling, could you make sure what I've said here
is accurate and/or relevant? I didn't know what else to say. ---AG}
The first of which was profiling with GHC.
We previously mentioned that profiling with GHC produced ugly code with
cost center annotations scattered throughout our code, but we also found it
very difficult to profile code at all.
In particular, profiling required that all installed libraries be re-compiled
with profiling support---which was a process that we could not replicate
consistently.

Another barrier we faced was how to test our code using QuickCheck.
Indeed, the barrier was so formidable in our case that we still have not been 
able to use it.
The problems we faced specifically were two-fold: How do we start testing
a complex module? and How can we use QuickCheck to generate random inputs with 
complex constraints?

Finally, one last barrier we faced was debugging runtime errors.
Runtime errors in the form of out-of-bounds indexing were particularly 
difficult to track down because they were not accompanied by a stack trace.
(While profiling with GHC could help remedy this, profiling was itself a 
barrier, as previous mentioned.)
We had to resort to polluting our code with cals to the \texttt{trace} function 
in order to track down where the out-of-bounds errors were occurring.

% - profiling 
% - debugging 
 % [ talk about profiling and debugging problems where?] 
% [ something about quickcheck: how do we know when we have enough QC props? How do we start testing a complex module? ]
